Can We Crunch the Social Graph on the Cheap?

### 2011-01-22 doc/hotpar2011/evaluation.tex
edits, rewriting, images

To study the memory concurrency in the system, we did experiments that approximate large irregular applications. The experiments consist of multiple threads each walking a number of separate linked lists. Each memory reference of a node is highly likely to result in an access to main memory, and we control the number of concurrent requests offered by a thread by varying the number of lists each thread walks. \todo{Brandon: somewhere we might explicitly say that this experiment involves no synchronization, and relate to plan to build this as mentioned in the components of runtime section} \todo{Brandon: is it important to mention that for isolation we use huge pages to reduce TLB misses?} Our experiments were run on a Xeon X5660, which uses the Nehalem microarchitecture, with 2 sockets, 6 cores per socket, and 12GB of DRAM per socket. For the results listed, memory allocations were pinned to socket0's DRAM.
\subsection{Results}

\todo{crop the pdfs}
\todo{enlarge font of tableau plot since pdf is getting scaled down}
\begin{figure*}
	\begin{center}
 		\includegraphics[width=0.5\textwidth]{figures/multi-listwalk-maxrates.pdf}
	\end{center}
	\caption{The max rates for processing list nodes with varying numbers of cores. Each core walks one list, and each node access results in a reference to main memory. All memory is local to socket0. Cores 1-6 are on socket0, and cores 7-12 are on socket1.}
	\label{fig:listwalk-maxrates}
\end{figure*}

\begin{figure*}
	\begin{center}
 		\includegraphics[width=0.5\textwidth]{figures/multi-listwalk-outflats.pdf}
	\end{center}
	\caption{}
	\label{fig:listwalk-outflats}
\end{figure*}

\todo{choose best phrase to use for outstanding/concurrent miss/refs/reads and use consistently unless there is a place one makes most sense}
\todo{if mentioning the cores 7-12 then need to talk about QPI link probably}
	 Figure~\ref{fig:listwalk-outflats} shows the total number of outstanding misses the system can handle before performance plateaus. We found that a single core is bandwidth-limited by the size of the L1 cache line-fill buffer (keeps track of outstanding misses of the core). Bandwidth \todo{word:performance?} plateaus at about 10 concurrent misses, the size of the buffer. Running multiple cores, the system reaches a bandwidth plateau with fewer outstanding misses per core. In fact, the maximum number of total memory references in flight on a single socket is consistently 36 (see figure). When using all six cores on the socket, bandwidth peaks at about 275Mref/s. Since this is less than even a conservative bound on available pin bandwidth, there is some bottleneck to performance in the memory system. A likely place is the global read-tracker queue containing references to the L3 cache; it has a capacity of 32 references, which concurs with the observed limit of 36 in-flight references. In addition, we performed measurements using performance counters, which consistently showed an average queue occupancy of 27, also close to the capacity 32.
 \todo{Brandon: it is great to identify why the bandwidth is limited, but in paper need to connect to why it is important. in general, we get an idea of what might need to be changed, especially as latency goes up and we need more concurrency}  \todo{Brandon: may cite nehalem bandwidth numbers here as well to validate our bw numbers with a second source}
	Bandwidth of memory requests from cores on a single socket seems to peak at about 275Mreq/s; however adding requests to socket0's memory from socket1 cores increases the bandwidth to about	350Mreq/s, showing that there is additional room in the memory system that is simply reserved for remote requests.\todo{Currently a bit of a 
hanging statement, but it is intended to show more evidence of limit being queue.}

@@

\subsection{Using green threads}
	If a core is to generate some number of concurrent memory requests to tolerate latency and keep busy, it may need multiple contexts \todo{Brandon: this idea likely goes elsewhere in describing overall approach}.  We ran our experiments with varying numbers of green threads per core to use multiple contexts to achieve concurrency. Our results, displayed in Figure~\ref{fig:listwalk-green-threads} \todo{decide whether max/avg rate as well as how much info/whether to make consistent with the other plots format}, show that using green threads gives little to no overhead due to context switching. \todo{Brandon: somehow also tie in prefetch-and-switch; also might elaborate on the results and yield-only if space}

>>>>

To study memory concurrency in commodity systems and begin building a runtime, we did experiments that approximate large irregular applications. The experiments consist of multiple threads each walking a number of separate linked lists. Each memory reference of a node is highly likely to result in an access to main memory, and we control the number of concurrent requests offered by a thread by varying the number of lists each thread walks. \todo{Brandon: somewhere we might explicitly say that this experiment involves no synchronization on list nodes, and relate to plan to build this as mentioned in the components of runtime section} Our experiments were run on a Xeon X5660, which uses the Nehalem microarchitecture, and has 2 sockets, 6 cores per socket, and 12GB of DRAM per socket. For the results listed, memory allocations were pinned to socket0's DRAM. Memory references from socket1's cores must be satisfied by sending requests over the QPI link between the sockets. \todo{Brandon: is it important to mention that for isolation we use huge pages to reduce TLB misses?}
\subsection{Results}

\todo{Might still consider the reqrate v. conc miss graph, because of showing the relation of performance going up as there is more concurrency, or else make sure to describe this clearly in the text}.
\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/multi-listwalk-maxrates-edited.pdf}
	\end{center}
	\caption{The max rates for processing list nodes with varying numbers of cores. Each core walks one list, and each node access results in a reference to main memory. All memory is local to socket0. Cores 1-6 are on socket0, and cores 7-12 are on socket1.}
	\label{fig:listwalk-maxrates}
\end{figure}

\begin{figure}[h]
	\begin{center}
		\includegraphics[width=0.5\textwidth]{figures/multi-listwalk-outflats-edited.pdf}
	\end{center}
	\caption{This is my caption about multi-listwak-outflats-edited, it shows the number of misses when rate plateaus...}
	\label{fig:listwalk-outflats}
\end{figure}

\todo{choose best phrase to use for outstanding/concurrent miss/refs/reads and use consistently unless there is a place one makes most sense}
	 Figure~\ref{fig:listwalk-outflats} shows the total number of outstanding misses the system can handle before performance plateaus. We found that for a single core, bandwidth plateaus once there are 10 concurrent misses; this is because the L1 cache line-buffer, which keeps track of outstanding misses of the core, is size 10. If concurrency is limited, there is less latency tolerance and performance is limited.

	Running multiple cores, the system reaches a bandwidth plateau with fewer outstanding misses per core. In fact, the maximum number of total memory references in flight on a single socket is consistently 36. When using all six cores on the socket, bandwidth peaks at about 275Mref/s (see Figure~\ref{fig:listwalk-maxrates}). Since this is less than even a conservative bound on available pin bandwidth to DRAM, there is some bottleneck to performance in the memory system. A likely place is the global read-tracker queue containing references to the L3 cache; it has a capacity of 32 references, which concurs with the observed limit of 36 in-flight references. In addition, we performed measurements using performance counters, which consistently showed an average queue occupancy of 27, also close to the capacity 32.
 \todo{Brandon: it is great to identify why the bandwidth is limited, but in paper need to connect to why it is important. in general, we get an idea of what might need to be changed, especially as latency goes up and we need more concurrency}  \todo{cite ``multi-socket memory concurrency" to validate bandwidth numbers}
	Bandwidth of memory requests from cores on a single socket seems to peak at about 275Mreq/s; however adding requests to socket0's memory from socket1 cores increases the bandwidth to about 350Mreq/s, showing that there is additional room in the memory system that is simply reserved for remote requests.\todo{Currently a bit of a 
hanging statement, but it is intended to show more evidence of limit being queue.}

@@

\subsection{Using green threads}
	If a core is to have the number of concurrent memory requests required to tolerate latency and keep busy, it may need multiple contexts \todo{Brandon: this idea likely goes elsewhere in describing overall approach}.  We ran our experiments with varying numbers of green threads per core to use multiple contexts to achieve concurrency. Our results, displayed in Figure~\ref{fig:listwalk-green-threads} \todo{decide whether max/avg rate as well as how much info/whether to make consistent with the other plots format}, show that using green threads gives little to no overhead due to context switching. \todo{Brandon: somehow also tie in prefetch-and-switch; also might elaborate on the results and yield-only if space}
